\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-14 paper/main.tex
fix sine curve plot annotation

When combining cilia motion and cardiac pulsations, we recognize flow features
persisting throughout the cardiac cycle, now with vortex structures in
the anterior and middle ventricles and directional flow in the
posterior ventricle (\Crefpanels{fig:fig1}{h--j}).
These characteristic patterns suggest that the flow in
the anterior and middle ventricles is primarily driven by the motile
cilia, while the cardiac pressure pulsations dominate within the
posterior ventricle and in the ducts connecting the ventricular
compartments. The CSF flow speed peaks at the time of peak caudal
cardiac pulsatile flow, when both the tangential traction induced by
the cilia and the pulsatile pressure-driven flow are aligned with the
rostrocaudal axis, reaching up to 27.9 \textmu m/s dorsally in the
posterior part of the middle ventricle. The flow Reynolds number is
$\mathrm{Re} \approx 0.004$, justifying the assumption of Stokes
flow. In summary, the computational CSF flow model reproduces the flow
features observed in vivo~\cite{Olstad2019CiliaryDevelopment}.

@@

To validate the model predictions, we performed photoconversion
experiments in transgenic zebrafish larvae expressing the secreted
Dendra2 in their brain ventricles. The approach is inspired by earlier work
where a photoconvertible protein was injected into the ventricle~\cite{fame2016directional}.
By using a transgenic system where the protein is directly secreted into the CSF,
we avoid detrimental effects of intraventricular injection on CSF properties. 
Our photoconversion protocol
consisted of the acquisition of a baseline fluorescence intensity
$F_0$, before localized exposure to a UV laser (at ROI 1) that
converts Dendra2 from a green- to a red-emitting fluorescent
protein. To quantify the transport of photoconverted proteins, we
measured and calculated the change in fluorescence intensity $\Delta F(t)
= (F(t)-F_0)/F_0$ over time (\Crefpanels{fig:fig3}{a}).

@@

\subsection{Conclusion and outlook}
Altogether, we have shown that the flow
generated by ciliary motion in the zebrafish larval brain ventricles
is of the scale of the geometry~\cite{Olstad2019CiliaryDevelopment}.
In contrast, one can expect that in large brains such as the
human brain, ciliary beating contributes only to flow proximal
to the ventricular walls\cite{Siyahhan2014FlowVentricles}.
Nevertheless, the cilia may still play an important role in regulating the
amount of substances, especially extracellular vesicles,
close to the ventricular walls,
similar to coral reefs where cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Notably, the discovery of a cilia-driven flow network
in rodent brains strongly supports such selective
brain ventricular transport~\cite{Faubel2016Cilia-basedVentricles}.
Our results also revealed that solute concentrations are affected
by ventricular geometry changes.
In fact, a 9\% difference in volume
had a noticeable impact on solute transport,
suggesting that pathological ventricular volume changes,
such as the enlargement observed in hydrocephalus,
could affect CSF distribution and composition.
Because CSF plays an important role in the regulation of solutes
within the ventricular system~\cite{yamada2021cerebrospinal},
understanding more of the interplay between
transport dynamics and ventricular morphology,
as well as the impact of
reduced mixing resulting from cilia-deficiency,
are important considerations for future work.

@@

% Data Availability Statement
\medskip
\noindent \textbf{Data Availability Statement} \par
\noindent The software and data used and presented in this paper are
openly available\cite{Herlyng2025Zenodo}.

% Conflic of Interest Disclosure
\medskip
\noindent \textbf{Conflict of Interest Disclosure} \par
\noindent The authors declare no conflicts of interest.

% Bibliography
\bibliographystyle{MSP}
\bibliography{ref}

\newpage 
%%--------------APPENDIX--------------%%
\appendix
\section{Supplementary Theory}
\subsection{Weak formulation of the Stokes equations}\label{subsec:appendixA1}
We consider the time-dependent Stokes equations
\begin{subequations}
\begin{alignat}{2}
   \rho\pdifft{\uu} - \nabla \cdot \bsig(\uu, p)
   &= \mathbf{0} \quad &&\mathrm{in} \ \Omega\times(0, T], \label{eq:BDM_stokes_mom}\\
  \nabla \cdot \uu &= 0 \quad &&\mathrm{in} \ \Omega\times(0, T], \label{eq:BDM_stokes_div} 
\end{alignat}
\label{eq:appendix_stokes_eqs}% 
\end{subequations}%
in the domain $\Omega$ with boundary $\Gamma$, where
$\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ is the viscous
stress tensor, in which $\beps(\uu) = \frac{1}{2}(\nabla\uu + (\nabla\uu)^T)$
is the symmetric gradient. Furthermore, $\uu\in\VV$ is the fluid velocity
and $p\in Q$ is the fluid pressure, and $\rho$ and $\mu$ are the fluid density
and dynamic viscosity, respectively. Unless otherwise stated, we let $\bsig = \bsig(\uu, p)$.

>>>>

When combining cilia motion and cardiac pulsations, we recognize flow features
persisting throughout the cardiac cycle, now with vortex structures in
the anterior and middle ventricles and directional flow in the
posterior ventricle (\Crefpanels{fig:fig1}{h--j},
\href{https://doi.org/10.5281/zenodo.15194757}{Supplementary Video 1}).
These characteristic patterns suggest that the flow in
the anterior and middle ventricles is primarily driven by the motile
cilia, while the cardiac pressure pulsations dominate within the
posterior ventricle and in the ducts connecting the ventricular
compartments. The CSF flow speed peaks at the time of peak caudal
cardiac pulsatile flow, when both the tangential traction induced by
the cilia and the pulsatile pressure-driven flow are aligned with the
rostrocaudal axis, reaching up to 27.9 \textmu m/s dorsally in the
posterior part of the middle ventricle. The flow Reynolds number is
$\mathrm{Re} \approx 0.004$, justifying the assumption of Stokes
flow. In summary, the computational CSF flow model reproduces the flow
features observed in vivo~\cite{Olstad2019CiliaryDevelopment}.

@@

To validate the model predictions, we performed photoconversion
experiments in transgenic zebrafish larvae expressing the secreted
Dendra2 in their brain ventricles. The approach is inspired by earlier work
where a photoconvertible protein was injected into the ventricle~\cite{fame2016directional}.
By using a transgenic system where the protein is directly secreted into the CSF,
we avoid detrimental effects of intraventricular injection on CSF properties. 
Our photoconversion protocol
consisted of the acquisition of a baseline fluorescence intensity
$F_0$, before localized exposure to a UV laser (at ROI 1) that
converts Dendra2 from a green- to a red-emitting fluorescent
protein (\href{https://doi.org/10.5281/zenodo.15194757}{Supplementary Video 2}).
To quantify the transport of photoconverted proteins, we
measured and calculated the change in fluorescence intensity $\Delta F(t)
= (F(t)-F_0)/F_0$ over time (\Crefpanels{fig:fig3}{a}).

@@

\subsection{Outlook}
Altogether, we have shown that the flow
generated by ciliary motion in the zebrafish larval brain ventricles
is of the scale of the geometry~\cite{Olstad2019CiliaryDevelopment}.
In contrast, one can expect that in large brains such as the
human brain, ciliary beating contributes only to flow proximal
to the ventricular walls\cite{Siyahhan2014FlowVentricles}.
Nevertheless, the cilia may still play an important role in regulating the
amount of substances, especially extracellular vesicles,
close to the ventricular walls,
similar to coral reefs where cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Notably, the discovery of a cilia-driven flow network
in rodent brains strongly supports such selective
brain ventricular transport~\cite{Faubel2016Cilia-basedVentricles}.
Our results also revealed that solute concentrations are affected
by ventricular geometry changes.
In fact, a 9\% difference in volume
had a noticeable impact on solute transport,
suggesting that pathological ventricular volume changes,
such as the enlargement observed in hydrocephalus,
could affect CSF distribution and composition.
Because CSF plays an important role in the regulation of solutes
within the ventricular system~\cite{yamada2021cerebrospinal},
understanding more of the interplay between
transport dynamics and ventricular morphology,
as well as the impact of
reduced mixing resulting from cilia-deficiency,
are important considerations for future work.

@@

% Data Availability Statement
\medskip
\noindent \textbf{Data availability statement} \par
\noindent The software and data used and presented in this paper are
openly available\cite{Herlyng2025Zenodo}.

% Conflic of Interest Disclosure
\medskip
\noindent \textbf{Conflict of interest disclosure} \par
\noindent The authors declare no conflicts of interest.

% Bibliography
\bibliographystyle{ieeetr}
\bibliography{ref}

\newpage 
%%--------------APPENDIX--------------%%
\appendix
\section{Supplementary theory and numerical methods}
\subsection{Weak formulation of the Stokes equations}\label{subsec:appendixA1}
We consider the time-dependent Stokes equations
\begin{subequations}
\begin{alignat}{2}
   \rho\pdifft{\uu} - \nabla \cdot \bsig(\uu, p)
   &= \mathbf{0} \quad &&\mathrm{in} \ \Omega\times(0, T], \label{eq:BDM_stokes_mom}\\
  \nabla \cdot \uu &= 0 \quad &&\mathrm{in} \ \Omega\times(0, T], \label{eq:BDM_stokes_div} 
\end{alignat}
\label{eq:appendix_stokes_eqs}% 
\end{subequations}%
in the domain $\Omega$ with boundary $\Gamma$, where
$\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ is the viscous
stress tensor, in which $\beps(\uu) = \frac{1}{2}(\nabla\uu + (\nabla\uu)^T)$
is the symmetric gradient. Furthermore, $\uu\in\VV$ is the fluid velocity
and $p\in Q$ is the fluid pressure, and $\rho$ and $\mu$ are the fluid density
and dynamic viscosity, respectively. Unless otherwise stated, we let $\bsig = \bsig(\uu, p)$.
